% !TeX root = ecn-encap-test_tr
% ================================================================
\section{Introduction}\label{ecn-encap-test_Intr}

This memo defines a brief set of tests to determine the decapsulation behaviour
of an unknown remote tunnel endpoint, with respect to the ECN field in the IP
header. It provides a table that says whether each possible detected behaviour
will propagate ECN correctly.

Test prerequisites are given in \S\,\ref{ecn-encap-test_Prereq}, the main hurdle
being the ability to overwrite the ECN field in the outer header at some point
along the span of a tunnel. This makes it hard to test `bump in the wire'
tunnels. To overcome this hurdle, a convenient arrangement would be to set up
the ingress of the tunnel under test on a host under the control of the tester.

The tests could be automated to be used by a tunnel ingress to determine whether
the egress it is paired with will propagate ECN correctly. Without such a test,
a tunnel ingress is required to zero the outer ECN field if it does not know
whether the egress it is paired with will propagate ECN
correctly~\cite{Briscoe16b:ecn-tunnel-scope_ID}.

In scenarios where there is no control protocol for a tunnel ingress to discover
the ECN capability of the egress, such a test could widen ECN coverage to
tunnelled paths where it is currently absent.

\subsection{Terminology}\label{ecn-encap-test_Term}

\begin{description}
	\item[Encap:] the encapsulation function at the ingress tunnel endpoint;
	\item[Decap:] The decapsulation function at the egress tunnel endpoint;
\end{description}

The following terms will be used for the IP header at different locations on the path relative to the tunnel, considering only the direction from application client to application server:
\begin{description}
	\item[Initial:] the header arriving at the tunnel ingress;
	\item[Inner:] the header that is encapsulated between the tunnel ingress and egress;
	\item[Outer:] the header that encapsulates the inner between the tunnel ingress and egress;
	\item[Onward:] the header leaving the tunnel egress.
\end{description}
